Epistemic Semantics in Guarded String Models

### 2020-07-18 tex/draft.tex
Different versions of Figure 4.
Edits up to fig. 4 in draft.text

@@

\section{Introduction and Related Work}
Linguistic semantics in the Montague tradition proceeds by assigning
propositional {\em semantic values} to disambiguated sentences of a
natural language.  A proposition is a set or class of {\em possible
  worlds}.  These are often assumed to be things with the same
nature and complexity as the world we occupy \citep{lewis1986plurality}.  But
alternatively, one can work with small idealized models, in order to
illustrate and test ideas.  To build such models, spaces of worlds and
individuals are stipulated as small finite sets, and semantic values
of lexical items are constructed as functions or relations from these
small sets.  Such toy or idealized models are useful in research and
in teaching, in that it is possible to represent propositions finitely
and explicitly, and to calculate with them. The point of this paper is
to scale up toy or idealized models to countable sets of worlds, and
to constructive and computable modeling of epistemic alternatives for
agents.  We describe a certain systematic way of defining such models,
and illustrate how to apply them in natural language semantics.
The focus is on epistemic semantics and clausal embedding. The fundamental move is to
identify possible worlds with strings of primitive events, so that
propositions are sets of strings. An advantage in this is that it
allows for a mathematical description of an algebra of a propositions,
coupled with a computational representation using either lazy lists of
strings, or finite state machines that describe sets of strings.

The approach taken here synthesizes five antecedents in a certain way.
John McCarthy's {\em Situation Calculus} is the source of the idea of
constructing possible worlds as event sequences
\citep{mccarthy1963situations,reiter2001knowledge}.  The algebraic
theory of {\em Kleene Algebra with Tests} characterizes algebras with
elements corresponding to propositions and event types in our
application \citep{kozen2001automata}.\marginpar{2001 or 2003?}  The
models we propose are an epistemic extension of guarded string models
for KAT, where a unary operation interpreted as an existential
epistemic modality is included for each agent.  {\em Action models} in
dynamic epistemic semantics introduced the technique of constructing
epistemic models from primitive alternative relations on events, in
order to capture the epistemic consequences of perceptual events
\citep{baltag1999logic}. This is the basis for our construction of
epistemic alternative relations.  Literature on finite state methods
in linguistic semantics has used event strings and sets of event
strings to theorize about tense and aspect in natural language
semantics
\citep{fernando2004finite,fernando2007observing,carlson2009tense} and
to express intensionality \cite{fernando2017intensions}.  Literature
on finite state intensiononal semantics has investigated how to do the
semantics of intensional complementation, including indirect
questions, in a setting where compositional semantics is expressed in
a finite state calculus \citep{rooth2017finite,collard2018finite}.  We
adopt this in our syntax-semantics interface for English.

@@

\begin{example}  \label{coin1}
\hspace{-4mm} \begin{tabular}[t]{ll}
  $a_1$  & Amy peeks at heads, by tipping the
  \\ & cup. Bob sees she's peeking, but not\\
  & what she sees. \\
  $b_1$  & Bob peeks at heads. \\ 
\end{tabular}
\end{example}

@@

In the discussion at the end of the last section, a sequence such as
$\bar{q}{p}$ can be viewed as a valuation of primitive test
propositions, which is used to describe world state. The primitives
are listed in fixed order, and left unmarked (indicating true) or
marked with the overbar (indicating false).  Suppose that in the coin
example, we have an additional primitive stative proposition (or
atomic test) $t$, interpreted as tails.  Since a coin is heads or
tails but not both, we want to allow the valuations $h\bar{t}$ and
$\bar{h}t$, and disallow $ht$ and $\bar{h}\bar{t}$.  This is enforced
by a {\em state formula}, which is a simply Boolean formula, in this
case \exi{effect}{a}.  Where {\sf{B}} is a set of atomic tests and
$\phi$ is a state constraint over {\sf{B}},
${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\phi}$ is the set of valuations of
{\sf{B}} that make formula $\phi$ true. The valuations are called
atoms, because they correspond to the atoms of Boolean algebra of
tests \citep{kozen2001automata}.

@@

Figure \ref{epikfigure} shows an Epik program that describes a
possible worlds model for two agents with information about one coin,
and events of the agents semi-privately looking at the coin. The line
beginning with {\tt state} enumerates ${\sf{B}}$. The
line beginning with {\tt constraint} gives the state formula.
The lines
beginning with {\tt event} declare events and their effect formulas.
Finally
the lines beginning with {\tt agent} define {\em event alternative}
relations for agents.  Each clause with an arrow has a single event
symbol on the left, and a disjunction of alternative events on the
right of the arrow. The interpretation of Amy's alternatives for $b_1$
(Bill peeks at heads), is that when $b_1$ happens, for Amy either
$b_1$ or $b_0$ (Bill peeks at tails) could be happening.

@@

Kleene Algebra with Tests is an algebraic theory that is defined by
equations and inequalities, which has model classes including
guarded string models, relational models, finite models, and matrix
models.  This paper focuses on defining a a concrete guarded
string algebra, the elements of which are sets of guarded strings, from
an Epik specification.
Definitions and notation mostly follow \cite{kozen2001automata}.
Additional syntax and semantics is included to cover multi-agent
epistemic semantics.  Guarded strings over a finite alphabet ${\sf{P}}$
are like ordinary strings, but with atoms 
alternating with the symbols
from ${\sf{P}}$.  In the algebra described by Figure \ref{epikfigure}, 
is the set of events $\{a_1,a_0,b_1,b_0\}$, and in the elevator example,
$\sf{P}$ is $\{u,d\}$.  In the elevator example, $\sf{B}$ is $\{p,q\}$, and in the coin
example it is $\{h,t\}$.
Assuming a trivially true state formula $\rho$,
${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\rho}$
in the elevator example is
$\{\bar q \bar p$, $\bar q p$, $q \bar p$, $q p\}$,
which we write $\{\hat0$, $\hat1$,
$\hat2$, $\hat3\}$.
In the coin example, as we already saw in
\ex{effect},
$\mathcal{A}^\varphi_{\mathsf B}$ is $\{h\bar{t},\bar{h}t\}$,
for which we use the shorthand $\{H,T\}$.
Guarded strings are strings of events, alternating with atoms,
and beginning and ending with atoms.
\ex{gs1} gives the encoding as guarded strings of the worlds in
\ex{elevator1} and \ex{coin1}.
The length of a guarded string
$p$, written $|p|$ is the number of events in $p$.
An atom such as $H$ is a guarded string of length $0$.

@@

Rather than individual guarded strings, elements of a guarded string
model for KAT are sets of guarded strings.  In the application, these
elements have the interpretation of propositions (sets of possible
worlds).  In a free guarded string model for KAT, any event can be
adjacent to any atom in a guarded string that is an element of the
underlying set for the algebra.  We instead impose the constraints
coming from the state and effect formulas.  \ex{wf1} defines the
well-formed guarded strings determined by and Epic specification.  (i)
says that each atom is consistent with the state constraint, and (ii)
says that each constituent token event $\alpha_ie_i\alpha_{i+1}$ is
consistent with the effect constraint on $e_i$.\footnote{
An alternative is to define equations such as
$\bar{\phi}=0$ (from the state formula $\phi$)
and $a_1\,{=}\,ha_{1}h$ (from the effect formula $h\mathbin{:}h$ for event $a_1$),
and construct a quotient algebra from the equivalence relation generated
by these equations.  This results in equating sets of guarded strings in the free algebra that
differ by guarded strings that are ill-formed according to the state and
effect formulas.  In the development in the text, we instead use
a set of guarded strings that are well-formed according to the state
and effect formulas as the representative of the equivalence class.
}

@@

Well-formed guarded strings have the interpretation of worlds in the
application to natural-language semantics.  The set of possible worlds
in the Kripke frame determined by an Epik specification is the set of
well-formed guarded strings.  At this point, we could say in the way
familiar from the type theory of possible worlds semantics that any
set of worlds is a proposition, so that the set of propositions is the
power set of the set of worlds \citep{montague,gallin}.  We will
instead define a more restrictive set of propositions corresponding to the regular sets
of strings. This is deferred to the next section.

In the natural language application, sets of well-formed guarded strings 
have the addional interpretation of event types.
An event-type is something that can ``happen'' is different worlds.
The event of the elevator going up is modeled not as the bare event
symbol $u$ or its unit set $\{u\}$, but as the set of guarded strings
\( \hat 0 \,u\, \hat 1, \hat 1 \,u\, \hat 2,  \hat 2 \,u\, \hat 3\).
The event of the elevator going up in a given world $w$ corresponds
to $w$ being incremented to form $w \diamond x$, where $x$ is an element
of the event type.  \ex{events1} gives the event types for
the examples.

@@

The construction so far defines a set of worlds $W$ from an Epik specification.
Normally and in our examples, $W$ is countably infinite, though some choices of effect
formulas can result in a finite set of worlds.
The next step is to define an alternative relation $R_a$ on worlds for each
agent $a$. This will result in a Kripke frame $\tuple{W,R_a,...}$ in the 
standard sense, consisting of a set of worlds, and a world-alternative relation
for each agent \citep{kripke, hintikka}.
An Epik specification defines an alternative relation on bare events for each
agent $a$, which we notate as $\hat{R}_{a}$.  This should be lifted to a relation
$R_a$ on worlds.  The rough idea is that when a world $w$ is incremented with
an event $e$, in the resulting world $w \diamond e$, epistemic alternatives
for agent $a$ are of the form $w' \diamond e'$, where $w'$ is an alternative to
for $a$ in $w$, and $e'$ is and event-alternative for $a$ (Moore --, Baltag --).\footnote{
 In this it is important that the event-alternative relation for an agent
 is constant across worlds. We anticipate that the definition given here produces
 results equivalent to what is found in literature on event alternatives in
 dynamic epistemic semantics, though we have not verified this.  That literature primarily focuses on mapping
 an epistemic model for a single time and situation to another, and uses general first-order models, rather than
 guarded string models. See ---, and articles in ---.  We picked up the idea from papers by Moss and his colleagues, together
 with Moore () and subsequent literature in situation theory.  This previous literature is motivated by epistemic
 logic and AI planning, rather than natural language semantics as in our application.
} This need to be implemented in a way that takes account of pre- and post-conditions for
events. For this, our approach is to refer the definition of well-formed guarded strings.
\ex{raise1} defines a relation on worlds from a relation on bare events.

\begin{example}  \label{raise1}
Let a set of guarded strings $W$ over events P and primitive tests B, and a relation $\hat{R}$ on P be given.
The corresponding relation $R$ holds between a guarded string
\[
\alpha_0e_0 ... e_n\alpha_{n+1} 
\]
in $W$ and a guarded string  $q$ iff $q$ is an element of $W$ and is of the form

@@

where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{\hat{R}}$.

\end{example}

This requires that in the alternative world $q$, each constituent event $e'_i$ is an
alternative to the event $e_i$ in the base world according to $\hat{R}$.
Compatibilities between events in the alternative world are enforced by the
requirement that the alternative $q$ is an element of $W$, so that 
state and effect formulas are enforced. Consider a scenario like the one from
Figure 1, with an additional agent Cem. The base world $T b_0 c_0$ is one where the
coin is tails, and first Bob looks at tails, and them Cem looks at tails. The first
event $b_0$ has the alternatives $b_0$ and $b_1$ for Aly, and the second event $c_0$
has the alternatives $c_0$ and $c_1$ for Aly.  This results in the combinations
$b_0c_0$, $b_0c_1$, $b_1c_0$, and $b_1c_1$. But these are filtered by post- and
pre-conditions of events in the alternative world, so that the set of alernatives for
Aly in $T b_0 c_0$ is $\{T b_0 c_0,H b_1 c_1\}$, with two world-alternatives instead
of four.

\section{The language of Epistemic KAT}
In addition to a set of possible worlds and an algebra of propositions, the
application in natural language requires a logical language for naming propositions.
The standard language for Kleene algebra with tests has the signature
$\tuple{K,+,\cdot,*,\bar{},0,1}$ \citep{kozen2001automata}.  In a guarded string
model for KAT, $K$ is a set of guarded strings, $+$ is set union, $\cdot$ is fusion
product raised to sets, $*$ is Kleene star, $\bar{}$ is complement for tests, $0$ the
empty set (functioning as an identity for $0$, and as False for tests) and 1 is the
set of atoms (functioning as the identity element for $\cdot$, and as True for
tests).  To this signature we add a unary modal operation $\Diamond_a$ for each
agent, and a unary complement operation $\not$ on elements of $K$. The latter is
included because natural languages have sentence negation, and it should be possible
to represent the propositional complement of a formula that captures the semantics of
a natural language sentence.  In addition, universal box modalities are defined as
duals of existential diamond modalities.

With modalities and propositional negation added, the signature of $n$-agent
epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1
  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language, but
using juxtaposition in place of the product symbol $\cdot$.
As illustrated in \ex{interpretation1}, terms in this language are used to
represent the proposositional semantic values of English sentences.\footnote{See Sections
6 and 7 for explanation and further examples.  }

\begin{example}   \label{interpretation1}
\begin{tabular}[t]{ll}
Bob knows that Amy knows whether it's tails.  \\
$\Box_b(\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t)$
\end{tabular}
\end{example}

\begin{figure}

  \[\begin{array}{l>{\hspace{-.5em}}c>{\hspace{-0.75em}}l}
  \multicolumn{3}{l}{\mbox{state terms $\sigma$ as in Figure 1}} \\
  \multicolumn 2 l {\textrm{events}} & e \in \mathsf{P} \\
  p,q & ::=& e~|~\sigma~|~p+q~|~pq~|~p^*~|~ \neg p ~|~ \Diamond_a\,p
  \end{array}\]

  Derived Operators\\[.1em]
  \(\Box_a p \triangleq \neg\Diamond_a\neg p\) \hfill \(p\&q \triangleq \neg(\neg p + \neg q)\) \hfill \(\_ \triangleq \sum\mathsf P\)

\caption{The language of Epik terms and key derived
  operators}
\label{fig:epik-grammar}
\label{termfigure}
\end{figure}

@@

\begin{figure}
  \[\begin{array}{l >{\triangleq}c l}
  \sem{ 0 }^{\epikspec} && \emptyset \\  
  \sem{ 1 }^{\varphi,{\mathcal{E}}} && \mathcal A^\varphi_{\mathsf B}\\
  \sem{ e }^{\varphi,{\mathcal{E}}}&& \hat{\mathcal E}^\varphi(e) \\
  \sem{ \psi }^{\varphi,{\mathcal{E}}} &&  \mathcal A^{\varphi~\psi}\\
  \sem{{p+q}}^{\varphi,{\mathcal{E}}} && \sem{{p}}^{\varphi,{\mathcal{E}}} \cup \sem{ q }^{\varphi,{\mathcal{E}}}\\
  \sem{{p;q}}^{\varphi,{\mathcal{E}}} && \sem{ p }^{\varphi,{\mathcal{E}}} \diamond \sem{{q}}^{\varphi,{\mathcal{E}}}\\
  \sem{ p^*}^{\varphi,{\mathcal{E}}} && \displaystyle\bigcup_n {\underbrace{\sem{ p }^{\varphi,{\mathcal{E}}}\diamond \ldots \diamond \sem{ p }^{\varphi,{\mathcal{E}}}}_{n}} \\
  \sem{ \neg p }^{\varphi,{\mathcal{E}}} && \sem{ \_^*}^{\varphi,{\mathcal{E}}} \setminus \sem{ p }^{\varphi,{\mathcal{E}}} \\
  \sem{ \Diamond_a~p } && \{x \mid x \mathbin{\hat{R_a}} y, y \in \sem{ p }^{\varphi,\mathcal E}\}

>>>>

\pagestyle{plain}

@@

\section{Introduction and Related Work}
Linguistic semantics in the Montague tradition proceeds by assigning
propositional {\em semantic values} to disambiguated sentences of a
natural language.  A proposition is a set or class of {\em possible
  worlds}.  These are often assumed to be things with the same
nature and complexity as the world we occupy \citep{lewis1986plurality}.  But
alternatively, one can work with small idealized models, in order to
illustrate and test ideas.  To build such models, spaces of worlds and
individuals are stipulated as small finite sets, and semantic values
of lexical items are constructed as functions or relations from these
small sets.  Such toy or idealized models are useful in research and
in teaching, in that it is possible to represent propositions finitely
and explicitly, and to calculate with them. The point of this paper is
to scale up toy or idealized models to countable sets of worlds, and
to constructive and computable modeling of epistemic alternatives for
agents.  We describe a certain systematic way of defining such models,
and illustrate how to apply them in natural language semantics.
The focus is on epistemic semantics and clausal embedding. The fundamental move is to
identify possible worlds with strings of primitive events, so that
propositions are sets of strings. An advantage in this is that it
allows for a mathematical description of an algebra of propositions,
coupled with a computational representation using either lazy lists of
strings, or finite state machines that describe sets of strings.

The approach taken here synthesizes five antecedents in a certain way.
John McCarthy's {\em Situation Calculus} is the source of the idea of
constructing possible worlds as event sequences
\citep{mccarthy1963situations,reiter2001knowledge}.  The algebraic
theory of {\em Kleene Algebra with Tests} characterizes algebras with
elements corresponding to propositions and event types in our
application \citep{kozen2001automata}.\marginpar{2001 or 2003?}  The
models we propose are an epistemic extension of guarded string models
for KAT, where a unary operation interpreted as an existential
epistemic modality is included for each agent.  {\em Action models} in
dynamic epistemic semantics introduce the technique of constructing
epistemic models from primitive alternative relations on events, in
order to capture the epistemic consequences of perceptual and
communicative events \citep{baltag1999logic}. This is the basis for
our construction of epistemic alternative relations.  Literature on
finite state methods in linguistic semantics has used event strings
and sets of event strings to theorize about tense and aspect in
natural language semantics
\citep{fernando2004finite,fernando2007observing,carlson2009tense} and
to express intensionality \cite{fernando2017intensions}.  Work on
finite state intensiononal semantics investigates how to do the
semantics of intensional complementation, including indirect
questions, in a setting where compositional semantics is expressed in
a finite state calculus \citep{rooth2017finite,collard2018finite}.  We
adopt this in our syntax-semantics interface for English.

@@

\begin{example}  \label{coin1}
\hspace{-4mm} \begin{tabular}[t]{ll}
  $a_1$  & Amy peeks at heads, by tipping the
  \\ & cup. Bob sees she's peeking, but not\\
  & what she sees. \\
  $b_1$  & Bob peeks at heads. \\
  $a_0$  & Amy peeks at tails. \\
  $b_0$  & Bob peeks at tails. \\     
\end{tabular}
\end{example}

@@

In the discussion at the end of the last section, a sequence such as
$\bar{q}{p}$ can be viewed as a valuation of primitive test
propositions, which is used to describe world state. The primitives
are listed in fixed order, and left unmarked (indicating true) or
marked with the overbar (indicating false).  Suppose that in the coin
example, we have an additional primitive stative proposition (or
atomic test) $t$, interpreted as tails.  Since a coin is heads or
tails but not both, we want to allow the valuations $h\bar{t}$ and
$\bar{h}t$, and disallow $ht$ and $\bar{h}\bar{t}$.  This is enforced
by a {\em state formula}, which is a Boolean formula, in this case the
one given on the second line of \ex{effect}.  Where {\sf{B}} is a set
of atomic tests and $\phi$ is a state constraint over {\sf{B}},
${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\phi}$ is the set of valuations of
{\sf{B}} that make formula $\phi$ true. Valuations are called atoms,
because they correspond to the atoms of a Boolean algebra of tests
\citep{kozen2001automata}.

@@

Epik is a specification language for possible worlds models that
includes declarations of events and states, state formulas, effect
formulas, and additional information.  Figure \ref{epikfigure} shows
an Epik program that describes a possible worlds model for two agents
with information about one coin, and events of the agents
semi-privately looking at the coin. The line beginning with {\tt
  state} enumerates ${\sf{B}}$. The line beginning with {\tt
  constraint} gives the state formula.  The lines beginning with {\tt
  event} declare events and their effect formulas.  Finally the lines
beginning with {\tt agent} define {\em event alternative} relations
for agents.  Each clause with an arrow has a single event symbol on
the left, and a disjunction of alternative events on the right of the
arrow. The interpretation of Amy's alternatives for $b_1$ (Bill peeks
at heads), is that when $b_1$ happens, for Amy either $b_1$ or $b_0$
(Bill peeks at tails) could be happening.

@@

This paper focuses on defining a concrete possible worlds model from
an Epik specification.  The models are an extension of guarded-string
models for Kleene Algebra with Tests (KAT). This is an algebraic
theory that has model classes including guarded string models,
relational models, finite models, and matrix models.  Our definitions
and notation follow \cite{kozen2001automata}.  We add syntax and
semantics is included to cover multi-agent epistemic semantics.

Guarded strings over a finite alphabet ${\sf{P}}$
are like ordinary strings, but with atoms over a set ${\sf{B}}$  
alternating with the symbols
from ${\sf{P}}$.  In the algebra described by Figure \ref{epikfigure}, 
${\sf{P}}$ is the set of events $\{a_1,a_0,b_1,b_0\}$, and
${\sf{B}}$ is $\{h,t\}$.
In the elevator example,
$\sf{P}$ is $\{u,d\}$, and $\sf{B}$ is $\{p,q\}$

Assuming a trivially true state formula $\rho$, the set of atoms
${\cal{A}}_{\mbox{\sf\tiny{B}}}^{\rho}$ in the elevator example is
$\{\bar q \bar p$, $\bar q p$, $q \bar p$, $q p\}$, which we write
$\{\hat0$, $\hat1$, $\hat2$, $\hat3\}$.  In the coin example, as we
already saw in \ex{effect}, $\mathcal{A}^\varphi_{\mathsf B}$ is
$\{h\bar{t},\bar{h}t\}$, for which we use the shorthand $\{H,T\}$.  A
guarded string over ${\sf{P}}$ and ${\sf{B}}$ is a strings of events
from ${\sf{P}}$, alternating with atoms over ${\sf{B}}$, and beginning
and ending with atoms.  \ex{gs1} gives the encoding as guarded strings
of the worlds in \ex{elevator1} and \ex{coin1}.  The length of a
guarded string $p$, written $|p|$ is the number of events in $p$.  An
atom such as $H$ is a guarded string of length $0$.

@@

Rather than individual guarded strings, elements of a guarded string
model for KAT are sets of guarded strings.  In our application, these
elements have the interpretation of propositions, which are sets of possible
worlds.  In a free guarded string model for KAT, any event can be
adjacent to any atom in a guarded string that is an element of the
underlying set for the algebra.  We instead impose the constraints
coming from the state and effect formulas.  \ex{wf1} defines the
well-formed guarded strings determined by and Epik specification.  Condition (i)
says that each atom is consistent with the state constraint, and condition (ii)
says that each constituent token event $\alpha_ie_i\alpha_{i+1}$ is
consistent with the effect constraint on $e_i$.\footnote{
An alternative is to define equations such as
$\bar{\phi}=0$ (from the state formula $\phi$)
and $a_1\,{=}\,ha_{1}h$ (from the effect formula $h\mathbin{:}h$ for event $a_1$),
and construct a quotient algebra from the equivalence relation generated
by these equations.  This results in equating sets of guarded strings in the free algebra that
differ by guarded strings that are ill-formed according to the state and
effect formulas.  In the development in the text, we instead use
a set of guarded strings that are well-formed according to the state
and effect formulas as the representative of the equivalence class.
}

@@

Well-formed guarded strings have the interpretation of worlds in the
application to natural-language semantics.  The set of possible worlds
in the Kripke frame determined by an Epik specification is the set of
well-formed guarded strings.  At this point, we could say that any set
of worlds is a proposition, so that the set of propositions is the
power set of the set of worlds
\citep{montague1975formal,gallin1975intensional}.  We will instead
define a more restrictive set of propositions corresponding to the
regular sets of strings. This is deferred to the next section.

In the natural language application, sets of well-formed guarded strings 
have the addional interpretation of event types.
An event-type is something that can ``happen'' is different worlds.
The event of the elevator going up is modeled not as the bare event
symbol $u$ or its unit set $\{u\}$, but as the set of guarded strings
\(\{ \hat 0 \,u\, \hat 1, \hat 1 \,u\, \hat 2,  \hat 2 \,u\, \hat 3 \}\).
The event of the elevator going up in a given world $w$ corresponds
to $w$ being incremented to form $w \diamond e$, where $e$ is an element
of the event type.  \ex{events1} gives the event types in
the examples.

@@

The construction so far defines a set of worlds from an Epik specification.  Normally
and in our examples, $W$ is countably infinite, though some choices of effect
formulas can result in a finite set of worlds.  The next step is to define an
alternative relation $R_a$ on worlds for each agent $a$. This will result in a Kripke
frame $\tuple{W,R_1,...,R_n}$ consisting of a set of worlds, and a world-alternative
relation for each agent \citep{kripke1963semantical}.  An Epik specification defines
an alternative relation on bare events for each agent $a$, which we notate as
${R}_{a}$.  This should be lifted to a relation $\hat{R}_a$ on worlds.  The basic
idea is that when a world $w$ is incremented with an event $e$, in the resulting
world $w \diamond e$, epistemic alternatives for agent $a$ are of the form $w'
\diamond e'$, where $w'$ is an alternative to for $a$ in $w$, and $e'$ is and
event-alternative to $e$ for $a$.\footnote{In this it is important that the
  event-alternative relation for an agent is constant across worlds. We anticipate
  that the definition given here produces results equivalent to what is found in
  literature on event alternatives in dynamic epistemic semantics, though we have not
  verified this.  That literature primarily focuses on mapping an epistemic model for
  a single time and situation to another, and uses general first-order models, rather
  than guarded string models. See \citet{baltag1999logic}, \citet{van2007dynamic}, and
  articles in \citet{ditmarsch2015handbook}.  We picked up the idea from papers by
  Moss and his colleagues, together with \citet{moore1984formal} and subsequent
  literature in situation theory.  This previous literature is motivated by epistemic
  logic and AI planning, rather than natural language semantics.  } This needs to be
implemented in a way that takes account of pre- and post-conditions for events. For
this, our approach is to refer the definition of well-formed guarded strings.
\ex{raise1} defines a relation on worlds from a relation on bare events.

\begin{example}  \label{raise1}
  Let $W$ be a set of guarded strings over  events P and primitive tests B, a relation ${R}$ be a relation
  on $P$. The corresponding relation $\hat{R}$ on $W$ holds between a guarded string
\[
\alpha_0e_0 ... e_n\alpha_{n+1} 
\]
in $W$ and a guarded string  $q$ iff $q$ is an element of $W$ and is of the form

@@

where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{\hat{R}}$.
\end{example}

This requires that in the alternative world $q$, each constituent event $e'_i$ is an
alternative to the event $e_i$ in the base world according to $\hat{R}$.
Compatibilities between events in the alternative world are enforced by the
requirement that the alternative $q$ is an element of $W$, so that state and effect
formulas are enforced.

Consider a scenario like the one from Figure 1, but with an additional agent Cem. The
base world $T b_0 T c_0 T$ is one where the coin is tails, and first Bob looks at
tails, and them Cem looks at tails. The first event $b_0$ has the alternatives $b_0$
and $b_1$ for Amy, and the second event $c_0$ has the alternatives $c_0$ and $c_1$
for Amy.  This results in four combinations $b_0c_0$, $b_0c_1$, $b_1c_0$, and
$b_1c_1$. But these are filtered by post- and pre-conditions of events in the
alternative world, so that the set of alernatives for Amy in $T b_0 T c_0 T$ is $\{T
b_0 T c_0 T,H b_1 H c_1 H\}$, with two world-alternatives instead of four.

\section{The logical language of Epistemic KAT}
In addition to a set of possible worlds and an algebra of propositions, the
application in natural language requires a logical language for naming propositions.
The standard language for Kleene algebra with tests has the signature
$\tuple{K,+,\cdot,*,\bar{},0,1}$ \citep{kozen2001automata}.  In a guarded string
model for KAT, $K$ is a set of sets of guarded strings, $+$ is set union, the
operation $\cdot$ is fusion product raised to sets, $*$ is Kleene star, the operation
$\;\bar{}\;$ is complement for tests, $0$ the empty set, and $1$ is the set of
atoms.\footnote{$0$ has the dual role the identity for $+$ (union), and as False for
  operations on tests.  $1$ has the dual role of the identity for product (fusion
  product raised to sets), and True for tests.  }. To this signature we add a unary
modal operation $\Diamond_a$ for each agent, and a unary complement operation $\neg$
on elements of $K$.  The intended interpretation of $\Diamond_a p$ is the set of
worlds where proposition $p$ is epistemically possible for agent $a$.  Propositional
complement is included because natural languages have sentence negation.  In
addition, universal box modalities are defined as duals of existential diamond
modalities.

With modalities and propositional negation added, the signature of $n$-agent
epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1
  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
Juxtaposition is used in place of the product symbol $\cdot$.  Terms in this language
are used to represent the proposositional semantic values of English sentences.
\ex{interpretation1} gives some examples.  To explain the first one $\bullet$ as
defined in Figure \ref{fig:epik-grammar} is the disjunction of the primitive events.
Since a world is a well-formed sequence of events, $\bullet *$ is the set of worlds.
Multiplying by the state symbol $h$ in the term $\bullet \! * \! h$ has the effect of
conjoining $h$ with the atom at the end of the world.  So $\bullet\!*\!h$ is the set
of worlds where the coin is heads.\footnote{A mapping from English sentences to logical
  terms in epistemic KAT is presented in Section 6}.

\begin{example}   \label{interpretation1}
\begin{tabular}[t]{ll}
${\bullet}\!* t$ & It's tails. \\
$\bullet \!* h$ & It's heads. \\  
$\Box_a \bullet \!* h$  & Amy knows that it's heads. \\
\end{tabular}
\vspace{1mm}

\begin{tabular}[t]{l}
  $\Box_b(\Box_a\!\bullet\!* t + \Box_a\!\bullet \!* \neg t)$\\
 Bob knows that Amy knows whether \\ \hspace{3mm} it's tails.  \\
\end{tabular}
\end{example}

\begin{figure}
\begin{tabular}{ll}
events & $e \in \mathsf{P}$ \\
states & $\sigma$ as in Figure 1 \\    
$p,q$ & $::= e~|~\sigma~|~p+q~|~pq~|~p^*~|~ \neg p ~|~ \Diamond_a\,p$\\
\end{tabular}

\vspace{1mm}
\begin{tabular}{cl}
$\Box_a p$ & $\triangleq \neg\Diamond_a\neg p$ \\
\(p \wedge q\) &  \(\triangleq \neg(\neg p + \neg q)\) \\
$\bullet$ & \( \triangleq \sum_{e \in \mathsf P}e\)
\end{tabular}

\caption{The language of Epik terms and key derived
  operators.}
\label{fig:epik-grammar}
\label{termfigure}
\end{figure}

@@

\newcommand{\AphiB}{\mathcal A^\varphi_{\mathsf B}}

\begin{figure}
  \[\begin{array}{l >{\triangleq}c l}
  \sem{ 0 }^{\epikspec} && \emptyset \\  
  \sem{ 1 }^{\epikspec} && \mathcal A^\varphi_{\mathsf B}\\
  \sem{ a }^{\epikspec} && \mathcal A^{a\varphi}_{\mathsf B}\\
  \sem{ \bar{\sigma} }^{\epikspec} && \AphiB \setminus \sem{\sigma}^{\epikspec}\\
  \sem{ e }^{\epikspec}&&  \setabs{\alpha e \beta}{\alpha \zeta_e \beta} \\
  \sem{{p+q}}^{\epikspec} && \sem{{p}}^{\epikspec} \cup \sem{ q }^{\epikspec}\\
  \sem{{pq}}^{\epikspec} && \setabs{x \diamond y}{ \begin{array}{l}x \in \sem{p}^{\epikspec} \\ y \in \sem{q}^{\epikspec}\\\mbox{$x \diamond y$ is defined}\end{array}} \\
  \sem{ p^*}^{\varphi,{\mathcal{E}}} && \displaystyle\bigcup_n {\underbrace{\sem{ p }^{\varphi,{\mathcal{E}}}\diamond \ldots \diamond \sem{ p }^{\varphi,{\mathcal{E}}}}_{n}} \\
  \sem{ \neg p }^{\varphi,{\mathcal{E}}} && \sem{ \_^*}^{\varphi,{\mathcal{E}}} \setminus \sem{ p }^{\varphi,{\mathcal{E}}} \\
  \sem{ \Diamond_a~p } && \{x \mid x \mathbin{\hat{R_a}} y, y \in \sem{ p }^{\varphi,\mathcal E}\}
